|

### 2019-12-19 appendix/appendix.Rmd
revision main text

---
title: |
       | Supplementary Materials for "Bottom-Up Accountability and Public Service Provision: Evidence from a Field Experiment in Brazil"
author: 
- "Danilo Freire[^freire]"
- "Manoel Galdino[^galdino]"
- "Umberto Mignozzetti[^mignozzetti]"
date: \today
thanks: "We thank Fábio Barros, Guilherme Jardim Duarte, Stephen Herzog, and David Skarbek for their helpful comments. We are specially grateful to Catarina Roman and the staff of Transparência Brasil for their excellent research assistance. This research was pre-registered on the EGAP pre-registry tool (<https://egap.org/registration/4505>) and approved by the Brazilian Ministry of Education. The authors thank the 2016 Google Social Impact Challenge for financial support and declare there are no conflict of interest. Replication data and code are available at <http://github.com/umbertomignozzetti/tdp-accountability>."
abstract: "Does grassroots participation improve public service delivery? We study the effect of a mobile phone application that allows citizens to monitor school construction projects in 1020 Brazilian municipalities. The app provides a platform where users can submit photos of construction sites, consult independent engineers, and contact the mayor's office about project delays. Our results show that the app has a null impact on the school construction outcomes. Additionally, we find that politicians are unresponsive to individual requests. The results question the impact of local monitoring on public service performance and suggest that interventions targeted at other groups may produce better policy outcomes."
abstractspacing: double
keywords: accountability, Brazil, impact evaluation, state capacity, technology
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: false
    keep_tex: no
    template: template.latex
---

>>>>

---
title: |
       | Supplementary Materials for "Bottom-Up Accountability and Public Service Provision: Evidence from a Field Experiment in Brazil"
author: 
- "Danilo Freire[^freire]"
- "Manoel Galdino[^galdino]"
- "Umberto Mignozzetti[^mignozzetti]"
date: \today
thanks: "We thank Fábio Barros, Guilherme Jardim Duarte, Stephen Herzog, and David Skarbek for their helpful comments. We are specially grateful to Catarina Roman and the staff of Transparência Brasil for their excellent research assistance. This research was pre-registered on the EGAP pre-registry tool (<https://egap.org/registration/4505>) and approved by the Brazilian Ministry of Education. The authors thank the 2016 Google Social Impact Challenge for financial support and declare there are no conflict of interest. Replication data and code are available at <http://github.com/umbertomignozzetti/tdp-accountability>."
abstract: "Does local oversight improve public service delivery? We study the effect of a mobile phone application that allows citizens to monitor school construction projects in Brazilian municipalities. The app provides a platform where users can submit photos of construction sites, consult independent engineers, and contact the mayor's office about project delays. Our results show that the app has a null impact on the school construction indicators. Additionally, we find that politicians are unresponsive to individual requests. The results question the impact of local monitoring on public service performance and suggest that interventions targeted at other groups may produce better policy outcomes."
abstractspacing: double
keywords: accountability, Brazil, impact evaluation, state capacity, technology
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: false
    keep_tex: no
    template: template.latex
---

### 2019-12-19 article/article.Rmd
revision main text

---
title:  |
        | Bottom-Up Accountability and Public Service Provision: Evidence from a Field Experiment in Brazil
author:
- "Danilo Freire[^freire]"
- "Manoel Galdino[^galdino]"
- "Umberto Mignozzetti[^mignozzetti]"
date: \today
thanks: "We thank Fábio Barros, Guilherme Jardim Duarte, Stephen Herzog, and David Skarbek for their helpful comments. We are specially grateful to Catarina Roman and the staff of Transparência Brasil for their excellent research assistance. This research was pre-registered on the EGAP pre-registry tool (<https://egap.org/registration/4505>) and approved by the Brazilian Ministry of Education. The authors thank the 2016 Google Social Impact Challenge for financial support and declare that there are no conflict of interest. Replication data and code are available at <http://github.com/umbertomignozzetti/tdp-accountability>."
abstract: "Does ~~grassroots participation~~ **local oversight** improve public service delivery? We study the effect of a mobile phone application that allows citizens to monitor school construction projects in Brazilian municipalities. The app provides a platform where users can submit photos of construction sites, consult independent engineers, and contact the mayor's office about project delays. Our results show that the app has a null impact on the school construction indicators. Additionally, we find that politicians are unresponsive to individual requests. The results question the impact of local monitoring on public service performance and suggest that interventions targeted at other groups may produce better policy outcomes."
abstractspacing: double
keywords: accountability, Brazil, impact evaluation, state capacity, technology
fontsize: 12pt
margin:
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    template: article-template.latex
---

@@

However, accountability systems **may** take many forms. One promising model is
that of _bottom-up monitoring_, in which citizens receive information about the
shortcomings of a given project so they can evaluate and pressure
underperforming public officials [@kosack2014does; @molina2016community;
@raffler2018weakness]. Proponents argue that bottom-up accountability is
effective because: 1) constituents have first-hand information about the
outcomes of local policies; 2) citizens have incentives to attack corruption
that directly affects themselves; 3) policy-makers are sensitive to social
punishment from their own communities [@serra2011combining, 570]. In this
regard, bottom-up accountability offers a potential solution to the
principal-agent dilemma in public service by aligning the interests of state
officials with those of the constituency they serve [@barro1973control;
@raffler2018weakness, 2].

@@

We use the TDP app to conduct two experimental interventions and test ~~its
impact~~ **the impact of citizen oversight** on five outcomes related to school
completion rates and complaints to public authorities. Overall, our results
show that providing information to citizens has no consistent impact on policy
outcomes. In the first experiment, we find that the TDP app increased the
likelihood of construction cancellation by 2 percent, but the result does not
replicate. The remaining five models have null results. In our second
intervention, none of the estimations reach conventional levels of statistical
significance. Importantly, all coefficients are small, what suggests that even
if the TDP app had a significant effect on the outcomes, its substantive impact
would be ~~virtually~~ negligible.

The findings raise questions about the ability of citizens to hold
representatives accountable using bottom-up ~~strategies~~ **monitoring**.
Although studies like @bjorkman2009power, @fiala2017social, and
@reinikka2005fighting report better policy outcomes after providing information
to local communities, our findings are closer to @banerjee2010pitfalls,
@keefer2014mass, @lieberman2014does, @bjorkman2017experimental,
@olken2007monitoring, and @raffler2018weakness, offering little evidence that
information-based interventions lead to greater government responsiveness. Our
analysis suggests that grassroots monitoring is ineffective in altering
government behaviour in Brazil.

@@

The _Tá de Pé_[^tdp] (TDP) cell phone application is an initiative carried out
by Transparência Brasil to foster bottom-up accountability in the Brazilian
public sector. More specifically, the main goal of the TDP project is to
improve responsiveness in government education expenditures. The TDP app
incentivises citizens to provide up-to-date information about unfinished school
constructions in their neighbourhoods, and that information will be assessed by
a group of independent specialists. In the case the construction is behind
schedule, TDP provides a writing platform whereby citizens can report to public
officials quickly and anonymously. The app then prompts citizens to write
a notification to the mayor's office, which has 15 days to reply. If they do
not respond to the request, the app forwards the notification to the Brazilian
Ministry of Education, making it harder for the municipality to access federal
funds in the future. The motivation behind this intervention is that providing
information to citizens empowers individuals to closely monitor public works.
This, in turn, results in better social outcomes as public agents become more
responsive to community demands.

@@

Between August 2017 and February 2019, we implemented two interventions to
measure the effect of the TDP app on five school construction outcomes plus
a placebo test. The outcomes are: 1) a placebo outcome indicating the
percentage of the project completed before the impact evaluation started; 2)
the percentage of the project completed by the end of the interventions; 3) the
difference between the percentage reported as completed before and after the
interventions; 4) the number of finished constructions; 5) the number of
cancelled constructions; 6) the number of schools where construction companies
updated the conclusion dates. In terms of expectations, Table
\ref{tab:outcomes} depicts the expected effects for each of the studied
outcomes[^more-details].

@@

The first intervention was carried out from August 2017 to February 2018 using
the Android version of TDP. The randomization was conduct at the municipal
level. We randomly selected 
`r as.numeric(table(impact_evaluation_phase1$treat)[1])` municipalities to the
control group and included 
`r as.numeric(table(impact_evaluation_phase1$treat)[2])` in the treatment
group. Our control condition consists in removing all information about school
construction from the TDP app in the chosen municipalities, so that citizens
were unable to report constructions in the control municipalities.

@@

We also conducted two manipulation checks and analysed the number of TDP app
downloads by municipality and over time. Figure \ref{fig:manipulation1}
displays the results and indicates that the treatment has good territorial
variability. There are `r length(unique(ga$municipality[ga$phase1==1]))`
downloads in the 
`r format(length(unique(impact_evaluation_phase1$municipality[impact_evaluation_phase1$treat=='Treatment'])),
big.mark = ",")` municipalities in the treatment condition. Downloads peak
during the Facebook TDP campaign, launched in October 2017, then diminish in
the following months. **Overall, 
`r format(sum(ga$novos_usuarios[ga$phase1==1]), big.mark = ",")` users
downloaded the app in phase one.**

```{r man1, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, cache=TRUE, out.width='.49\\linewidth', fig.width=4, fig.height=3.5, fig.show='hold', fig.align='center', fig.cap='\\label{fig:manipulation1}Manipulation checks for intervention 1. The first plot shows the geographical distribution of the treatment condition and the second graph displays the number of TDP app downloads from August 2017 to February 2018.'}

@@

## Map GA intervention 1
mapbr <- get_brmap("City")
gaaux <- filter(ga, phase1 == 1)
gaaux <- select(gaaux, ibge_code7, downl)
gaaux <- rename(gaaux, City = ibge_code7)
gaaux <- unique(gaaux)
gaaux$city <- as.numeric(gaaux$City)
gaaux2 <- impact_evaluation_phase1 %>%
  select(ibge_code7) %>%
  transmute(City = as.numeric(ibge_code7)) %>%
  unique()
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 0
gaaux2 <- data.frame(City = mapbr$City)
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 2
gaaux$Downloads <- factor(gaaux$downl,
													labels = c("No Downloads", "Downloads", "Not in App"))
mapbr <- join_data(mapbr, gaaux, by = "City")
plot_downloads_int1 <- ggplot() +
  geom_sf(data = mapbr, aes(fill = Downloads), lwd = 0.01) +
  scale_fill_manual(values = c("#E69F00", "#0072B2", "#F0E442")) +
  theme_map()
plot_downloads_int1

@@

\newpage
The second intervention is similar to intervention 1 in all but three
characteristics. First, the TDP app was **then** available for both Android and
iOS devices. Second, we randomised the intervention at the school level, with
`r as.numeric(table(impact_evaluation_phase2$treat)[1])` control and 
`r format(as.numeric(table(impact_evaluation_phase2$treat)[2]), big.mark = ",",
scientific = FALSE)` treatment units. We used blocked randomisation stratified
by Brazilian states, school construction status (under construction, stopped,
unfinished), and whether the municipality spent more on school construction
that the distribution median. Finally, the intervention period lasted from
August 2018 to February 2019.

@@

```{r man2, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, cache=TRUE, out.width='.49\\linewidth', fig.width=4, fig.height=3.5, fig.show='hold', fig.align='center', fig.cap='\\label{fig:manipulation2}Manipulation checks for intervention 2. The graphs display the geographical and temporal variation of TDP app downloads from August 2018 to February 2019.'}
# Load data
load("tdp.RData")

@@

## Map manipulation test -- Intervention 2
mapbr <- get_brmap("City")
gaaux <- filter(ga, phase2 == 1)
gaaux <- select(gaaux, ibge_code7, downl)
gaaux <- rename(gaaux, City = ibge_code7)
gaaux <- unique(gaaux)
gaaux$City <- as.numeric(gaaux$City)
gaaux2 <- impact_evaluation_phase2 %>%
  select(ibge_code7) %>%
  transmute(City = as.numeric(ibge_code7)) %>%
  unique()
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 0
gaaux2 <- data.frame(City = mapbr$City)
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 2
gaaux$Downloads <- factor(gaaux$downl, labels = c("No Downloads", "Downloads",
                                                  "Not in App"))
mapbr <- join_data(mapbr, gaaux, by = "City")
p1 <- ggplot() +
  geom_sf(data = mapbr, aes(fill = Downloads), lwd = 0.01) +
  scale_fill_manual(values = c("#E69F00", "#0072B2", "#F0E442")) +
  theme_map()
p1

@@

We find that the app only has a small effect on cancellation rates. The TDP
application increases the likelihood of cancelling the construction in 
`r round(100*resInt1[[2]][[20]]$coefficients[1], 2)` percent. While this result
goes in the opposite direction of our theoretical expectations, the finding is
inconsistent and does not replicate in the second experiment. All other
coefficients are not statistically significant at conventional levels. On the
one hand, the results indicate that our placebo outcome, the percentage of the
invested executed before the intervention, indeed behaves as predicted. On the
other hand, we expected the five remaining outcomes to improve after the
introduction of the app. The literature on bottom-up accountability argues that
delivering more information about the shortcomings of public services provision
will ~~mobilise citizens who, in turn, would pressure state agents to produce
better social policies~~ **put citizens in a position where they can monitor
state agents and improve provider behaviour** [@raffler2018weakness, 2] . Our
results do not lend support to that hypothesis.

@@

In this paper, we discuss **whether delivering information to citizens via** a
mobile phone application fosters ~~grassroots mobilisation~~ **community
oversight** and political accountability in Brazil. Our two interventions show
that the results are at best mixed. Although we find ~~an~~ **some treatment**
effect on school construction completion and cancelling rates in the first
intervention, the app has no consistent impact on our outcomes of interest.
These findings add to the studies that cast doubts on the relationship between
bottom-up accountability and local policy performance [e.g.,
@banerjee2010pitfalls; @lieberman2014does; @raffler2018weakness].

@@

In sum, our experiments suggest that popular participation and bottom-up
monitoring may not be effective to improve public service delivery in the case
of school constructions in Brazil. Nevertheless, the null findings are
informative to researchers and policy-makers. The most important recommendation
derived from this study is that interventions targeting elite groups, such as
lobbyists or civil servants, might render better policy outcomes than those
focused at the grassroots levels. Another core lesson is that although digital
interventions are promising means to deliver information, perhaps they do not
have the same impact as personal, face-to-face communication. Since many
developing countries share Brazil's issues with education provision, the
shortcomings we describe here serve as warnings for future interventions.
Finally, whether this study generalises beyond school constructions to other
bottom-up programs, and to other contexts, remains to be studied.

>>>>

---
title:  |
        | Bottom-Up Accountability and Public Service Provision: Evidence from a Field Experiment in Brazil
author:
- "Danilo Freire[^freire]"
- "Manoel Galdino[^galdino]"
- "Umberto Mignozzetti[^mignozzetti]"
date: \today
thanks: "We thank Fábio Barros, Guilherme Jardim Duarte, Stephen Herzog, and David Skarbek for their helpful comments. We are specially grateful to Catarina Roman and the staff of Transparência Brasil for their excellent research assistance. This research was pre-registered on the EGAP pre-registry tool (<https://egap.org/registration/4505>) and approved by the Brazilian Ministry of Education. The authors thank the 2016 Google Social Impact Challenge for financial support and declare that there are no conflict of interest. Replication data and code are available at <http://github.com/umbertomignozzetti/tdp-accountability>."
abstract: "Does ~~grassroots participation~~ **local oversight** improve public service delivery? We study the effect of a mobile phone application that allows citizens to monitor school construction projects in Brazilian municipalities. ~~The app provides a platform where users can submit photos of construction sites, consult independent engineers, and contact the mayor's office about project delays.~~ **The app prompts users to submit data about construction sites, sends such crowdsourced information to independent engineers, and contacts the mayors' offices about project delays.** Our results show that the app has a null impact on the school construction indicators. Additionally, we find that politicians are unresponsive to individual requests. The results question the impact of ~~local~~ **bottom-up** monitoring on public service performance and suggest that interventions targeted ~~at other groups~~ **to other groups, or focusing on different issues,** may produce better policy outcomes."
abstractspacing: double
keywords: accountability, Brazil, impact evaluation, state capacity, technology
fontsize: 12pt
margin:
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    template: article-template.latex
---

@@

However, accountability ~~systems take many forms~~ **mechanisms extend beyond
elections.** One promising model is that of _bottom-up monitoring_, in which
citizens receive information about the shortcomings of a given project so they
can evaluate and pressure underperforming public officials [@kosack2014does;
@molina2016community; @raffler2018weakness]. Proponents argue that bottom-up
accountability is effective because: 1) constituents have first-hand
information about the outcomes of local policies; 2) citizens have incentives
to attack corruption that directly affects themselves; 3) policy-makers are
sensitive to social punishment from their own communities [@serra2011combining,
570]. In this regard, bottom-up accountability offers a potential solution to
the principal-agent dilemma in public service by aligning the interests of
state officials with those of the constituency they serve [@barro1973control;
@raffler2018weakness].

@@

We use the TDP app to conduct two experimental interventions and test ~~its
impact~~ **the impact of citizen oversight** on five outcomes related to school
completion rates and complaints to public authorities. Overall, ~~our results
show that~~ providing information to citizens has no consistent impact on
policy outcomes. In the first experiment, we find that the TDP app increased
the likelihood of construction cancellation by 2 percent, but the result does
not replicate. The remaining five models have null results. In our second
intervention, none of the estimations reach conventional levels of statistical
significance. Importantly, all coefficients are small, what suggests that even
if the TDP app had a significant effect on the outcomes, its substantive impact
would be ~~virtually~~ negligible.

The findings raise questions about the ability of citizens to hold
representatives accountable using bottom-up ~~strategies~~ **monitoring**. 
\sout{Although studies like} **On the one hand,** @bjorkman2009power,
@lagunes2018guardians, and @reinikka2005fighting report better policy outcomes
after providing information to local communities. ~~our findings are closer
to~~ **On the other hand,** @banerjee2010pitfalls, @keefer2014mass,
@lieberman2014does, @bjorkman2017experimental, @olken2007monitoring, and
@raffler2018weakness, find little evidence that information-based interventions
lead to greater government responsiveness. **Our results are in line with the
latter group and suggest that local oversight is ineffective in altering
government behaviour in Brazil.**

@@

The _Tá de Pé_[^tdp] (TDP) cell phone application is an initiative carried out
by Transparência Brasil to foster bottom-up accountability in the Brazilian
public sector. More specifically, the main goal of the TDP project is to
improve responsiveness in government education expenditures. The TDP app
incentivises citizens to provide up-to-date information about unfinished school
constructions in their neighbourhoods, and that information will be assessed by
a group of independent specialists. In the case the construction is behind
schedule, TDP provides a writing platform whereby citizens can report to public
officials quickly and anonymously. The app then writes a notification to the
mayor's office, which has 15 days to reply. If they do not respond to the
request, the app forwards the notification to the Brazilian Ministry of
Education, making it harder for the municipality to access federal funds in the
future. The motivation behind this intervention is that providing information
to citizens empowers individuals to closely monitor public works. This, in
turn, results in better social outcomes as public agents become more responsive
to community demands.

@@

Between August 2017 and July 2019, we implemented two interventions to measure
the effect of the TDP app on five school construction outcomes plus a placebo
test. The outcomes are: 1) a placebo outcome indicating the percentage of the
project completed before the impact evaluation started; 2) the percentage of
the project completed by the end of the interventions; 3) the difference
between the percentage reported as completed before and after the
interventions; 4) the number of finished constructions; 5) the number of
cancelled constructions; 6) the number of schools where construction companies
updated the conclusion dates. ~~In terms of expectations,~~ Table
\ref{tab:outcomes} depicts the expected effects for each of the studied
outcomes[^more-details].

@@

The first intervention was carried out from August 2017 to July 2018 using the
Android version of TDP. The randomization was conduct at the municipal level.
We randomly selected `r as.numeric(table(impact_evaluation_phase1$treat)[1])`
municipalities to the control group and included `r
format(as.numeric(table(impact_evaluation_phase1$treat)[2]), big.mark = ",")`
in the treatment group. Our control condition consists in removing all
information about school construction from the TDP app in the chosen
municipalities, so that citizens were unable to report constructions in the
control municipalities.

@@

We also conducted two manipulation checks and analysed the number of TDP app
downloads by municipality and over time. Figure \ref{fig:manipulation1}
displays the results and indicates that the treatment has good territorial
variability. There are `r length(unique(ga$municipality[ga$phase1==1]))`
downloads in the 
`r format(length(unique(impact_evaluation_phase1$municipality[impact_evaluation_phase1$treat=='Treatment'])),
big.mark = ",")` municipalities in the treatment condition. Downloads peak
during the Facebook TDP campaign, launched in October 2017, then diminish in
the following months. **Overall, 
`r format(sum(ga$novos_usuarios[ga$phase1==1]), big.mark = ",")` users
downloaded the app during intervention 1.**

```{r man1, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, cache=TRUE, out.width='.49\\linewidth', fig.width=4, fig.height=4, fig.show='hold', fig.align='center', fig.cap='\\label{fig:manipulation1}Manipulation checks for intervention 1. The first plot shows the geographical distribution of the treatment condition and the second graph displays the number of TDP app downloads from August 2017 to July 2018.'}

@@

## Map GA intervention 1
mapbr <- get_brmap("City")
gaaux <- filter(ga, phase1 == 1)
gaaux <- select(gaaux, ibge_code7, downl)
gaaux <- rename(gaaux, City = ibge_code7)
gaaux <- unique(gaaux)
gaaux$city <- as.numeric(gaaux$City)
gaaux2 <- impact_evaluation_phase1 %>%
  select(ibge_code7) %>%
  transmute(City = as.numeric(ibge_code7)) %>%
  unique()
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 0
gaaux2 <- data.frame(City = mapbr$City)
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 2
gaaux$Downloads <- factor(gaaux$downl,
													labels = c("No Downloads", "Downloads", "Not in App"))
mapbr <- join_data(mapbr, gaaux, by = "City")
plot_downloads_int1 <- ggplot() +
  geom_sf(data = mapbr, aes(fill = Downloads), lwd = 0.001) +
  scale_fill_manual(values = c("#E69F00", "#0072B2", "#F0E442")) +
  theme_map()
plot_downloads_int1

@@

The second intervention is similar to intervention 1 in all but three
characteristics. First, the TDP app was **then** available for both Android and
iOS devices. Second, we randomised the intervention at the school level, with
`r as.numeric(table(impact_evaluation_phase2$treat)[1])` control and 
`r format(as.numeric(table(impact_evaluation_phase2$treat)[2]), big.mark = ",",
scientific = FALSE)` treatment units. We used blocked randomisation stratified
by Brazilian states, school construction status (under construction, stopped,
unfinished), and whether the municipality spent more on school construction
that the distribution median. Finally, the intervention period lasted from
August 2018 to July 2019.

@@

```{r man2, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, cache=TRUE, out.width='.49\\linewidth', fig.width=4, fig.height=4, fig.show='hold', fig.align='center', fig.cap='\\label{fig:manipulation2}Manipulation checks for intervention 2. The graphs display the geographical and temporal variation of TDP app downloads from August 2018 to July 2019.'}
# Load data
load("tdp.RData")

@@

## Map manipulation test -- Intervention 2
mapbr <- get_brmap("City")
gaaux <- filter(ga, phase2 == 1)
gaaux <- select(gaaux, ibge_code7, downl)
gaaux <- rename(gaaux, City = ibge_code7)
gaaux <- unique(gaaux)
gaaux$City <- as.numeric(gaaux$City)
gaaux2 <- impact_evaluation_phase2 %>%
  select(ibge_code7) %>%
  transmute(City = as.numeric(ibge_code7)) %>%
  unique()
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 0
gaaux2 <- data.frame(City = mapbr$City)
gaaux <- merge(gaaux2, gaaux, all.x = TRUE); rm(gaaux2)
gaaux$downl[is.na(gaaux$downl)] <- 2
gaaux$Downloads <- factor(gaaux$downl, labels = c("No Downloads", "Downloads",
                                                  "Not in App"))
mapbr <- join_data(mapbr, gaaux, by = "City")
p1 <- ggplot() +
  geom_sf(data = mapbr, aes(fill = Downloads), lwd = 0.001) +
  scale_fill_manual(values = c("#E69F00", "#0072B2", "#F0E442")) +
  theme_map()
p1

@@

We find that the app only has a small effect on cancellation rates. The TDP
application increases the likelihood of cancelling the construction in 
`r round(100*resInt1[[2]][[20]]$coefficients[1], 2)` percent. While this result
goes in the opposite direction of our theoretical expectations, the finding is
inconsistent and does not replicate in the second experiment. All other
coefficients are not statistically significant at conventional levels. On the
one hand, the results indicate that our placebo outcome, the percentage of the
invested executed before the intervention, indeed behaves as predicted. On the
other hand, we expected the five remaining outcomes to improve after the
introduction of the app. The literature on bottom-up accountability argues that
delivering more information about the shortcomings of public services provision
will ~~mobilise citizens who, in turn, would pressure state agents to produce
better social policies~~ **put citizens in a position where they can monitor
state agents and improve provider behaviour** [@raffler2018weakness]. Our
results do not lend support to that hypothesis.

@@

In this paper, we discuss **whether delivering information to citizens via** a
mobile phone application fosters ~~grassroots mobilisation~~ **community
oversight** and political accountability in Brazil. Our two interventions show
that the results are at best mixed. Although we find ~~an~~ **some treatment**
effect on school cancelling rates in the first intervention, the app has no
consistent impact on our outcomes of interest. These findings add to the
studies that cast doubts on the relationship between bottom-up accountability
and local policy performance [e.g., @banerjee2010pitfalls; @lieberman2014does;
@raffler2018weakness].

@@

In sum, our experiments suggest that popular participation and bottom-up
monitoring may not be effective to improve public service delivery in the case
of school constructions in Brazil. Nevertheless, the null findings are
informative to researchers and policy-makers. The most important recommendation
derived from this study is that interventions targeting elite groups, such as
lobbyists or civil servants, might render better ~~policy~~ **school
construction** outcomes than those focused at the ~~grassroots levels~~
**community level**. Another core lesson is that although digital interventions
are promising means to deliver information, perhaps they do not have the same
impact as personal, face-to-face communication. Since many developing countries
share Brazil's issues with education provision, the shortcomings we describe
here serve as warnings for future interventions. Finally, whether this study
generalises beyond school constructions to other bottom-up programs, and to
other contexts, remains to be studied.

### 2019-12-19 article/letter.Rmd
revision main text

\noindent 2 December 2019

@@

\noindent Sincerely,

\noindent The Authors

@@

**1)** The editor writes:

\noindent 
**Response:**

\begin{quotation}
Lorem ipsum.
\end{quotation}

@@

**1)** Reviewer 1 writes:

\noindent
**Response:**

>>>>

\noindent 17 December 2019

@@

\vspace{.5cm} \noindent Sincerely,

\vspace{.5cm} \noindent The Authors

@@

**1)** The editor writes: "_The reviewer confirmed my own sense that the
manuscript and underlying field experiment are interesting, although you will
see that the reviewer suggests many ways the manuscript can be improved.  I am
in full agreement with the reviewer that you need to reflect more deeply about
the theory and evidence for why information provision via the app could be
connected to mobilization, -- otherwise the null result is not interesting.  As
the reviewer notes, the null result could be due to a disconnect between
information provision and action, or due to who selects into downloading the
app._"

\vspace{.5cm} \noindent \textbf{Response:}

\begin{quotation}
We agree with the editor and the reviewer in regards to the link between
information provision and mobilisation. We believe the main issue lies in the
lack of evidence for mobilisation, that is, information on how users interacted
with the application and whether they have used the app consistently. In the
revised version of the manuscript, we provide data from Google Analytics which
indicate that users did engage with the app. According to the service, the app
had 6,092 active users in intervention 1 and gained 4,078 new users during
intervention 2. On average, each user launched 60 sessions per app install, an
indication of their actual engagement with the application. Moreover, the data
also show that the app had more than 53,000 screen visualisations, with an
average of 2.42 visualisations per user. We have added this information to the
main text and the online appendix. The additions are marked in \textbf{boldface}.
We elaborate further on this issue in the response to Reviewer 1 below.
\end{quotation}

@@

**1)** Reviewer 1 writes: "_My biggest concern is there is no discussion of the
actual bottom-up accountability that may (or may not) have occurred. [...] The
authors have devised a nice experiment and I believe their design to have been
appropriately administrated. But the design hinges on several critical
mechanisms that go largely undiscussed: not only must information be available
to citizens such that they might participate in local oversight, but said
citizens must 1) access the information and 2) act on said information. This is
mentioned in the Discussion---that perhaps the app was insufficient to spur
collective action---but that also seems like an empirical question the authors
could feasibly address. [...] Once people downloaded the App, did they use
it?_"

\vspace{.5cm} \noindent \textbf{Response:}

\begin{quotation}
This first point raised by the reviewer is connected to the response we have
provided above. We have made several changes to the main paper and the appendix
to address this important concern. First, we have added a the following section
to the appendix in order to clarify how the TDP app works:

3. Treatment Definition and Mechanism

The \textit{Tá de Pé} app

1. Users download the app
2. Use the app and report the schools to the mayors' office
3. Our team took the reported schools and started implementing the pressure
on the mayors' office.
4. Should the pressure have work, we would witness improvements on governmental
	 data regarding the school construction progress.

For the step (4), we relied on monthly issued administrative data by the
Brazilian Ministry of Education. The Ministry of Education provides monthly
data on the school constructions that are being build with Federal resources,
and that are delayed across the country. To compute the delay we use the
finishing dates reported by the municipality at the time they sign the
agreement with the Federal Government. This data helped us to keep the App
updated and the outcomes in the paper are selected to be six months before
(pre-treatment) and after the intervention. This data was constantly updated by
the Ministry of Education, and we found no red flags in the overall quality of
the data.

Therefore, steps (1), (3), and (4) were standardized to not vary across the
country. The only source of variation in our intervention should come from step
(2).

In step (2), we applied the treatment by removing schools within municipalities
out of the app. In the first intervention, that happened at the municipal
level, we excluded all schools within a given randomly selected municipality.
In the second intervention, we excluded randomly selected schools out of the
App. Removing a school means that the school was in the dataset provided by the
Brazilian Ministry of Education, but it did not show up in the App. Therefore,
our treatment effect represents the school being present (treatment) or absent
(control) in the App.

In terms of what this represents, we argue that the App is promoting two
crucial features for bottom-up accountability: first, it is facilitating people
to connect with information that would be otherwise be in an obscure
spreadsheet at the Ministry of Education website; second, whenever people find
this information troublesome, we are providing an easy way to them to act upon
the information. As the design was intuitive and straightforward, we believe
that most of Brazilians would find it easy to use the App.

Moreover, the treatment to be successful does not require that everyone in the
municipality uses the App. Clearly that more usage means more pictures and
denounces, and this represents more pressure over the mayors office, what could
revert into a higher effect. However, just one denounce would be enough, as it
can be directed to the Brazilian General Comptroller (CGU), and this could
potentially deny future funding to the municipality.

\end{quotation}
